shrimp_goby_boatnoise

### 2025-02-25 ms_boatnoise_shrimp_goby.Rmd
Script tidy, PDF figures, and update ignore

**SUMMARY**

**AIM**

**AUTHORS**

**AFFILIATIONS**

**CONTRIBUTOR ROLES**

@@

@@

```{r, message=FALSE, results='hide'}
#install.packages("pacman")
#require("pacman") 
pacman::p_load(tidyverse, ggplot2, brms, cmdstanr, broom, broom.mixed, tidybayes, ggthemes, emmeans, ggdist, gridExtra, bayesplot, see, patchwork, ggtext, extraDistr, gghalves, ggpp, rstan, gt)

@@

@@

- **💿 BORIS_output.csv**: This is the summary output from the BORIS, the key logging software used for behavioural observations. This includes the occurrence count and total duration of each of the measured behaviour endpoints, for each shrimp and each goby across all trials. They are identified with the code column.  
- **💿 Video_codes.csv**: This has meta data for each video, include treatment information and associated codes.
- **💿 species.csv**: This is the species name for the shrimp in each of the videos

```{r}
#assigning the directory so that the code can be universally read on any computer as long as the entire project file is shared
directory <- './2-data/'

#reading in the coded raw BORIS output
BORIS_output <- read.csv("./2-data/BORIS_output.csv")

#reading in the decoder
Video_codes <- read.csv("./2-data/Video_codes.csv")

#reading in shrimp species information
shrimp_species <- read.csv("./2-data/species.csv")
```

@@

🚧 Jack, maybe you could break this up and add a little more detial, I got a little lost in the code 🚧

@@

Restructure the data frame for easier analysis

@@

@@

Run model

@@

@@

```{r}
goby_out_plot3
ggsave("./3-figs/goby_out_plot3.jpeg", plot = goby_out_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
```

@@

shrimp_out_fig_two
```
#### Combining plots into one figure

@@

shrimp_out_plot3
ggsave("./3-figs/shrimp_out_plot3.jpeg", plot = shrimp_out_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
```

@@

```{r}
shrimp_contact_plot3
ggsave("./3-figs/shrimp_contact_plot3.jpeg", plot = shrimp_contact_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
```

>>>>

**🚧 UNPUBLISHED AND ONGOING 🚧 **   
The data in this script is currently **unpublished**. There may be part of this script that are currently incomplete. This script has been made publicly available for easy of sharing with collaborators and colleagues who are interested in the progression of the project and for complete transparency.

-------

**SUMMARY**  
Human activities are reshaping marine soundscapes, with boat noise a near-constant source of underwater noise and a primary contributor to rising oceanic noise pollution. Despite increasing concern, empirical data for most marine life, particularly in natural settings, remains sparse. Fish and crustaceans are underrepresented in noise pollution research, despite their auditory ranges make them especially susceptible to low-frequency typical of boat engines. Accordingly, this study employs a field-based experimental approach to investigate the impacts of boat noise on an iconic behavioural mutualism between gobies (Amblyeleotris steinitzi) and shrimp (Alpheus spp). Our findings demonstrate that both species exhibit increased burrow use in response to boat noise, suggesting a heightened perception of risk. Despite the perceived risk triggered by boat noise, the behaviour of the two species remained tightly correlated and communication—measured as antennae contact—remained unchanged. Notably, shrimp exhibited stronger responses to 2-stroke engines, whereas gobies responded more strongly to 4-stroke engines, indicating species-specific sensitivities to different noise spectra. These results highlight the complex and taxon-specific effects of noise pollution on marine species and emphasize the need for targeted research to inform conservation strategies aimed at mitigating anthropogenic acoustic disturbances in marine ecosystems .

-------

**AIM**  
This study employs a field-based experimental approach to investigate the impacts of boat noise on an iconic behavioural mutualism between gobies (Amblyeleotris steinitzi) and shrimp (Alpheus spp). We test the hypothesis that boat noise exposure (2-stroke and 4-stroke engines) would alter the mutualism of these two species by independently altering perceived risk, leading to shifts in behaviour, and potentially disrupting their communication.

-------

**AUTHORS**  
Jack L. Manera^1†^   
Jake M. Martin^1,2,3†^    
Maria M. Palacios^4^ 
Rachel T Mason^3^    
Mark I. McCormick^5^    
Bob BM. Wong^1^

-------

**AFFILIATIONS**    
^1^School of Biological Sciences, Monash University, Clayton, Victoria 3800, Australia   
^2^School of Life and Environmental Sciences, Deakin University, Waurn Ponds, Victoria 3216, Australia   
^3^Department of Wildlife, Fish, and Environmental Studies, Swedish University of Agricultural Sciences, Umeå, Sweden   
^4^Centre for Nature Positive Solutions, School of Science, RMIT University, Melbourne, Victoria 3100, Australia   
^5^Coastal Marine Field Station, School of Science, University of Waikato, Tauranga, New Zealand   
^†^These authors contributed equally as lead investigators

-------

**CONTRIBUTOR ROLES**  
*🚧 to be added 🚧*

<!------------------------------->
# 📑 Sharing/accessing and citing
<!------------------------------->

**Licenses/restrictions placed on the data:** CC-BY 4.0

**Link to the associated publication:**     
🚧 *To be added* 🚧

**Recommended citation for this data:**      
🚧 *To be added based on OSF DOI* 🚧

@@

**Jack L. Manera**

📧 **Email:** [jack.manera@monash.edu](mailto:jack.manera@monash.edu)

🌐 **Web:** *🚧 to be added 🚧*

🐙 **GitHub**: [JLManera](https://github.com/JLManera)

-------

@@

```{r, message=FALSE, results='hide'}
#install.packages("pacman")
#install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos")))

pacman::p_load(tidyverse, ggplot2, brms, cmdstanr, broom, broom.mixed, tidybayes, ggthemes, emmeans, ggdist, gridExtra, bayesplot, see, patchwork, ggtext, extraDistr, gghalves, ggpp, rstan, gt)

@@

## Directories

Assigning the directory so that the code can be universally read on any computer as long as the entire project file is shared. This is a relative path.

```{r}
directory <- './2-data/'
```

@@

- **💿 BORIS_output.csv**: This is the summary output from the BORIS, the key logging software used for behavioural observations. This includes the occurrence count and total duration of each of the measured behaviour endpoints, for each shrimp and each goby across all trials. They are identified with the code column.

```{r}
BORIS_output <- read.csv("./2-data/BORIS_output.csv")
```

- **💿 Video_codes.csv**: This has meta data for each video, include treatment information and associated codes.

```{r}
Video_codes <- read.csv("./2-data/Video_codes.csv")
```

- **💿 species.csv**: This is the species name for the shrimp in each of the videos

```{r}
shrimp_species <- read.csv("./2-data/species.csv")
```

@@

🚧 Jack, maybe you could break this up and add a little more detail, I got a little lost in the code 🚧

@@

Restructure the data frame for easier analysis. We also write this file as **💿 final_df.csv** in the the 📁 `./2-data` folder. This is what we will use for analysis.

@@

### Final model structure

This is the structure of the final model

@@

@@

Here we reload the saved model.

@@

```{r}
goby_out_plot3
ggsave("./3-figs/goby_out_plot3.jpeg", plot = goby_out_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
ggsave("./3-figs/goby_out_plot3.pdf", plot = goby_out_plot3, units = "px", width = 2200, height = 1400)
```

@@

shrimp_out_fig_two
```

@@

shrimp_out_plot3
ggsave("./3-figs/shrimp_out_plot3.jpeg", plot = shrimp_out_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
ggsave("./3-figs/shrimp_out_plot3.pdf", plot = shrimp_out_plot3, units = "px", width = 2200, height = 1400)
```

@@

```{r}
shrimp_contact_plot3
ggsave("./3-figs/shrimp_contact_plot3.jpeg", plot = shrimp_contact_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
ggsave("./3-figs/shrimp_contact_plot3.pdf", plot = shrimp_contact_plot3, units = "px", width = 2200, height = 1400)
```
